---
title: "Unicorn companies: An R programing data analysis project"
output: 
  html_document:
    theme: 'simplex'
    toc: yes
---

# Introduction

------------------------------------------------------------------------

Are you ready to dive deeper into our analysis? We'll be taking a structured approach to ensure our examination is both accurate and precise. Our guide consists of six steps: Ask, Prepare, Process, Analyze, Visualize, and Share. Each stage has its unique advantages, allowing us to derive insights that can be trusted. By following this process, we'll gain a deeper understanding of the data and the story it tells. The data we have is as of **MARCH 2022** So let's get started and uncover the insights that await us!

# Ask

------------------------------------------------------------------------

The "Ask" phase is crucial for defining the scope and objectives of our analysis. Here are the questions that we need to answer to gain a deeper understanding of the unicorn landscape:

1.  Which are the top 10 countries with the highest unicorn valuations?

2.  What is the global distribution of Unicorns?

3.  What is the market share of total unicorn valuations by industry?

4.  What is the total unicorn valuation by industry?

5.  Is there a positive correlation between the total amount raised and the company's valuation?

6.  When were most of the unicorns founded, and when did they achieve unicorn status?

7.  How long did it take for unicorns to reach their unicorn status from the day they were founded?

# Prepare / Process

------------------------------------------------------------------------

In the "Prepare" phase, it's important to make sure that our data is clean and formatted correctly so that we can analyze it effectively. Using R programming language can be a powerful tool for data cleaning and manipulation. Let's get started and walk through the code chunks and their outputs to ensure we understand what's happening in each step.

#### **First, importing libraries for processing, analyzing and visualizing data**

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

# Inserting Libreries

library(tidyverse)
library(tmap)
library(tmaptools)
library(RColorBrewer)
library(plotly)

# Importing df 

unicorn_companies <- 
  as_tibble(
    read_csv("unicorn_companies_2022.csv"))

# Presenting df

library(knitr)
library(kableExtra)
library(htmltools)

kable(unicorn_companies, "html") %>%
  kable_styling(full_width = FALSE, 
                bootstrap_options = c(
                  "striped", "hover", "condensed", "responsive"), 
                fixed_thead = TRUE, font_size = 10) %>%
  scroll_box(height = "300px", width = "100%")

```

#### **Cleaning column names to match `snake_case` naming conventions**

```{r echo=FALSE}
colnames(unicorn_companies)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies <- 
  unicorn_companies %>% 
  rename_with(~ tolower(gsub(' ', '_', .x, fixed = TRUE))) %>% 
  rename('investors' = 'select_inverstors', 'valuation' ='valuation_($b)')

```

```{r echo=FALSE}
colnames(unicorn_companies)
```

#### **Cleaning duplicates based on 'company' column**

```{r echo=FALSE}
sum(duplicated(unicorn_companies$company))
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

unicorn_companies %>% 
  filter(duplicated(company))

unicorn_companies <- distinct(unicorn_companies, company, .keep_all = TRUE)

```

```{r echo=FALSE}
sum(duplicated(unicorn_companies$company))
```
#### **Cleaning `valuation` column by removing '\$' and changing data type**

```{r echo=FALSE}
head(unicorn_companies$valuation, 10)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies$valuation <- 
  gsub('\\$', '', unicorn_companies$valuation)

unicorn_companies$valuation <- 
  as.numeric(unicorn_companies$valuation)

```

```{r echo=FALSE}
head(unicorn_companies$valuation, 10)
```

#### **Cleaning `date_joined` column data type**

```{r echo=FALSE}
head(unicorn_companies$date_joined, 10)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies$date_joined <- 
  as.Date(
    unicorn_companies$date_joined, '%m/%d/%Y')

```

```{r echo=FALSE}
head(unicorn_companies$date_joined, 10)
```

#### **Cleaning `industry` column categorical values**

```{r echo=FALSE}
unique(unicorn_companies$industry)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies <- unicorn_companies %>% 
  mutate(
    industry = if_else(
      industry %in% c(
        'Vision Plus Capital, GSR Ventures, ZhenFund',
        'Vertex Ventures SE Asia, Global Founders Capital, Visa Ventures',
        'Tiger Global Management, Tiger Brokers, DCM Ventures',
        'Temasek, Guggenheim Investments, Qatar Investment Authority',
        'SingTel Innov8, Alpha JWC Ventures, Golden Gate Ventures',
        'Sequoia Capital, Thoma Bravo, Softbank',
        'Sequoia Capital China, Shunwei Capital Partners, Qualgro',
        'Sequoia Capital China, ING, Alibaba Entrepreneurs Fund',
        'Mundi Ventures, Doqling Capital Partners, Activant Capital',
        'Kuang-Chi',
        'Jungle Ventures, Accel, Venture Highway',
        'Hopu Investment Management, Boyu Capital, DC Thomson Ventures',
        'B Capital Group, Monk\'s Hill Ventures, Dynamic Parcel Distribution',
        'Andreessen Horowitz, DST Global, IDG Capital',
        '500 Global, Rakuten Ventures, Golden Gate Ventures',
        'Dragonfly Captial, Qiming Venture Partners, DST Global'
      ), 
      'Ventures and Funds', 
      industry
    ),
    industry = recode(
      industry, 
      'Finttech' = 'Fintech', 
      'Artificial intelligence' = 'Artificial Intelligence'
    )
  )

```

```{r echo=FALSE}
unique(unicorn_companies$industry)
```

#### **Cleaning `investors` column by replacing 'None' with 'NA'**

```{r}
nrow(unicorn_companies[unicorn_companies$investors == 'None', ])
```

```{r}
unicorn_companies$investors <- gsub('None', NA, unicorn_companies$investors)
```

#### **Cleaning outliers in `founded_year` column**

```{r echo=FALSE}
unicorn_companies %>%
  distinct(founded_year, .keep_all = TRUE) %>%
  arrange(founded_year) %>%
  na.omit() %>%
  pull(founded_year)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

unicorn_companies <- 
  unicorn_companies %>% 
  filter(founded_year >= 1990)

```

```{r echo=FALSE}
unicorn_companies %>%
  distinct(founded_year, .keep_all = TRUE) %>%
  arrange(founded_year) %>%
  na.omit() %>%
  pull(founded_year)
```

#### **Cleaning `founded_year` data type and replacing 'None' with 'NA'**

```{r echo=FALSE}
nrow(unicorn_companies[unicorn_companies$founded_year == 'None', ])
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

  # Replacing 'None' with 'NA'

unicorn_companies$founded_year <- gsub('None', NA, unicorn_companies$founded_year)

  # Data type to date

unicorn_companies$founded_year <- 
  as.Date(
    format(
      paste0('01/01/', unicorn_companies$founded_year)), 
    '%m/%d/%Y')

```

#### **Cleaning `total_raised` column by removing symbols and and replacing 'None' with 'NA'**

```{r echo=FALSE}
head(unicorn_companies$total_raised, 10)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

  # Extract the last character of the 'total_raised' col into a new col 'total_raised_ext'

unicorn_companies$total_raised_ext <- 
  substr(
    unicorn_companies$total_raised, 
    nchar(unicorn_companies$total_raised), 
    nchar(unicorn_companies$total_raised))

  # Remove unwanted char from the "total_raised" col

unicorn_companies$total_raised <- 
  gsub('\\$', '', unicorn_companies$total_raised)
unicorn_companies$total_raised <- 
  gsub('B', '', unicorn_companies$total_raised)
unicorn_companies$total_raised <- 
  gsub('M', '', unicorn_companies$total_raised)
unicorn_companies$total_raised <- 
  gsub('None', 'NA', unicorn_companies$total_raised)
unicorn_companies$total_raised <- 
  gsub('K', '', unicorn_companies$total_raised)

  # Convert the 'total_raised' col to numeric data type

unicorn_companies$total_raised <- 
  as.numeric(unicorn_companies$total_raised)

  # Multiply the 'total_raised' col by the factor in the 'total_raised_ext' col

for (i in 1:nrow(unicorn_companies)) {
  if (unicorn_companies$total_raised_ext[i] == 'B') {
    unicorn_companies$total_raised[i] <- unicorn_companies$total_raised[i] * 1000000000
  } else if (unicorn_companies$total_raised_ext[i] == 'M') {
    unicorn_companies$total_raised[i] <- unicorn_companies$total_raised[i] * 1000000
  } else if (unicorn_companies$total_raised_ext[i] == 'K') {
    unicorn_companies$total_raised[i] <- unicorn_companies$total_raised[i] * 1000
  }
}

remove(i)

  # Drop the 'total_raised_ext' col from the data frame

unicorn_companies$total_raised_ext <- NULL

```

```{r echo=FALSE}
head(unicorn_companies$total_raised, 10)
```

#### **Checking `financial_stage` / `investors_count` / `deal_terms` / `portfolio_exits` for values == 'None'**

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

nrow(unicorn_companies[unicorn_companies$financial_stage == 'None', ])

nrow(unicorn_companies[unicorn_companies$investors_count == 'None', ])

nrow(unicorn_companies[unicorn_companies$deal_terms == 'None', ])

nrow(unicorn_companies[unicorn_companies$portfolio_exits == 'None', ])

```

#### **Dropping `financial_stage` and `portfolio_exits` for high number of missing values**

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies$financial_stage <- NULL

unicorn_companies$portfolio_exits <- NULL
```

#### **Replacing `investors_count` and `deal_terms` None values with NA**

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies$investors_count <- gsub('None', NA, unicorn_companies$investors_count)

unicorn_companies$deal_terms <- gsub('None', NA, unicorn_companies$deal_terms)

```

#### **Cleaning variables data types**

```{r echo=FALSE}
sapply(unicorn_companies, class)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

unicorn_companies$country <- 
  as.factor(unicorn_companies$country)

unicorn_companies$city <- 
  as.factor(unicorn_companies$city)

unicorn_companies$industry <- 
  as.factor(unicorn_companies$industry)

unicorn_companies$investors_count <- 
  as.integer(unicorn_companies$investors_count)

unicorn_companies$deal_terms <- 
  as.integer(unicorn_companies$deal_terms)
```

```{r echo=FALSE}
sapply(unicorn_companies, class)
```

#### **Inspect the data after cleaning**

```{r}

kable(unicorn_companies, "html") %>%
  kable_styling(full_width = FALSE, 
                bootstrap_options = c(
                  "striped", "hover", "condensed", "responsive"), 
                fixed_thead = TRUE, font_size = 10) %>%
  scroll_box(height = "300px", width = "100%")

```

# Analyze / Visualize

------------------------------------------------------------------------

Lets play!

#### **Which are the top 10 countries with the highest unicorn valuations?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=10, fig.height=5, fig.align='center'}
unicorn_companies %>%
  group_by(country) %>%
  summarize(valuation_sum = sum(valuation)) %>%
  arrange(desc(valuation_sum)) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(country, -valuation_sum),
             y = valuation_sum,
             fill = valuation_sum)) +
  geom_bar(aes(x = reorder(country, -valuation_sum),
               y = valuation_sum,
               fill = valuation_sum),
           stat = "identity") +
  scale_fill_gradient(low = 'cadetblue2',
                      high = 'dodgerblue4') +
  labs(x = 'Country',
       y = 'Total valuation of unicorns ($B)',
       title = 'Valuation of unicorns in top 10 countries ($B)') +
  geom_text(aes(label = paste0('$ ', round(valuation_sum), ' B')),
            position = position_dodge(width=0.9),
            vjust=-0.5,
            size = 3) +
  theme_grey() +
  theme(plot.title = element_text(face = "bold"), 
        axis.text.x = element_text(angle = 45, size = 9),
        legend.position = 'none',
        panel.background = element_rect(fill = 'grey90'))

```

#### **What is the global distribution of Unicorns?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=9.5, fig.height=5, fig.align='center'}
map_data <- map_data("world") %>%
  mutate(region = gsub('USA', 'United States', region)) %>%
  mutate(region = gsub('UK', 'United Kingdom', region)) %>%
  left_join(
    unicorn_companies %>%
      group_by(country) %>%
      tally() %>% 
      arrange(desc(n)),
    by = c("region" = "country"))

ggplotly(
  ggplot(map_data, aes(x = long, y = lat, group = group)) + 
    geom_polygon(
      aes(
        fill = cut(n, c(0, 10, 50, 100, 500, Inf), 
                   labels = c("1-24", "25-49", "50-99", "100-499", "500+")), 
        text = paste(
          "<b>", region, "</b><br>", 
          "Unicorns = ", n, "<br>"
        )
      ),
      color = 'white'
    ) + 
    scale_fill_manual(
      values = c(
        "1-24" = '#ffa600',
        "25-49" = '#ff6361',
        "50-99" = '#bc5090',
        "100-499" = '#58508d',
        "500+" = '#003f5c'
      ),
      na.value = "grey90"
    ) +
    theme_void() +
    labs(x = 'Industry',
         y = 'Total valuation ($B)', 
         title = 'Unicorns world map by number of companies',
         fill = "Number of Unicorns") +
    theme(
      plot.title = element_text(face = "bold"), 
      axis.line = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank()
    ),
  tooltip = 'text'
) 

```

#### **What is the market share of total unicorn valuations by industry?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, , fig.width=10, fig.height=5, fig.align='center'}
industry_n <- 
  unicorn_companies %>%
  group_by(industry) %>% 
  summarise(valuation_industry = sum(valuation))

industry_n <- 
  industry_n %>% 
  mutate(
    percentage_valuation = round(
      industry_n$valuation_industry / sum(industry_n$valuation_industry)*100)
  )

industry_n$industry <- 
  reorder(industry_n$industry, industry_n$valuation_industry)

industry_n %>% 
  ggplot(
    aes(
      x="", 
      y = valuation_industry, 
      fill= industry)
  ) +
  geom_bar(
    width = 1, 
    stat="identity", 
    color = 'white'
  ) +
  geom_text(
    aes(
      x= 1.4, 
      label = paste0(percentage_valuation, ' %')
    ), 
    position = position_stack(vjust = 0.5), 
    size = 3
  ) +
  theme_classic() +
  coord_polar(
    "y", 
    start=0
  ) +
  scale_fill_discrete(
    limits=rev(levels(factor(industry_n$industry))), 
    name = 'Industry'
  ) + 
  theme(
    plot.title = element_text(face = "bold"),
    axis.line = element_blank(), 
    axis.text = element_blank(), 
    axis.ticks = element_blank()
  ) + 
  labs(
    x = NULL, 
    y = NULL, 
    fill = NULL, 
    title = 'Unicorns market share by industry'
  )

```

#### **What is the total unicorn valuation by industry?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=12, fig.height=5, fig.align='center'}
unicorn_companies %>%
  group_by(industry) %>%
  summarize(valuation_sum = sum(valuation)) %>%
  arrange(desc(valuation_sum)) %>%
  ggplot(
    aes(
      x = fct_reorder(industry, valuation_sum), 
      y = valuation_sum, 
      fill = valuation_sum
    )
  ) +
  geom_bar(
    stat = "identity"
  ) + 
  coord_flip() +
  scale_fill_gradient(
    low = 'cadetblue2', 
    high = 'dodgerblue4'
  ) +
  labs(
    x = 'Industry',
    y = 'Total valuation ($B)', 
    title = 'Unicorns market valuation by industry'
  ) + 
  geom_text(
    aes(
      label = paste0('$', round(valuation_sum), 'B')
    ), 
    vjust = 0.3, 
    hjust = -0.2, 
    size = 3
  ) + 
  theme_gray() +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = 'none'
  )
```

#### **Is there a positive correlation between the total amount raised and the company's valuation?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=9.5, fig.height=5, fig.align='center'}
 ggplotly(unicorn_companies %>% 
             ggplot(
               aes(
                 x = valuation,
                 y = total_raised / 1000000000, 
                 color = total_raised / 1000000000, 
                 size = valuation, 
                 label = company,
                 alpha = 0.7
               )
             ) + 
             geom_point(
               aes(
                 text = paste(
                   "<b>", company, "</b><br><br>", 
                   "Industry:", industry, "<br>", 
                   "Total raised = $", scales::comma(round(total_raised / 1000000000, 2)), "B<br>", 
                   "Valuation = $", scales::comma(round(valuation, 1)), "B"
                 )
               )
               ,alpha = 0.7
             ) + 
             scale_color_gradient(
               low = 'cadetblue2', 
               high = 'dodgerblue4'
             ) +
             scale_size_binned(
               range = c(0.1, 5)
             ) + 
             labs(
               title = 'Correlation between total raised and valuation', 
               color = 'Total raised ($B)',
               x = 'Total raised ($B)', 
               y = 'Valuation ($B)'
             ) + 
             theme_gray() +
             theme(plot.title = element_text(face = "bold")), 
           tooltip = 'text')

```

#### **When were most of the unicorns founded, and when did they achieve unicorn status?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=9.5, fig.height=5, fig.align='center'}
founded_joined <- 
  as_tibble(
    data.frame(
      date_joined = as.integer(format(as.Date(unicorn_companies$date_joined), "%Y")), 
      founded_year = as.integer(format(as.Date(unicorn_companies$founded_year), "%Y"))))

founded_joined$unicorn_time <- 
  founded_joined$date_joined - founded_joined$founded_year 

founded_n <- founded_joined %>% 
  group_by(founded_year) %>% 
  filter(founded_year >= 1985) %>% 
  tally()

joined_n <-  founded_joined %>% 
  group_by(date_joined) %>% 
  tally()

ggplotly(
  ggplot() + 
    geom_area(
      data = founded_n, 
      aes(
        x = founded_year, 
        y = n), 
      color = 'cadetblue3', 
      fill = 'lightcoral', 
      alpha = 0.6
      ) + 
    geom_point(
      data = founded_n, 
      aes(
        x = founded_year, 
        y = n, 
        text = paste(
          "<b>Founded year: ",founded_year, "</b><br>", 
          "Number:", n, "<br>")), 
      color = 'dodgerblue4', 
      size = 0.9) + 
    geom_area(
      data = joined_n, 
      aes(
        x = date_joined, 
        y = n), 
      color = 'firebrick', 
      fill = 'dodgerblue4', 
      alpha = 0.3
      ) + 
    geom_point(
      data = joined_n, 
      aes(
        x = date_joined, 
        y = n, 
        text = paste(
          "<b>Year joined: ",date_joined, "</b><br>", 
          "Number:", n, "<br>")), 
      color = 'firebrick4', 
      size = 1
      ) + 
    scale_x_continuous(
      breaks = seq(1900, 2030, by = 10)) + 
    labs(title = 'Companies by year founded vs year joined unicorn club', 
         x = 'Years', 
         y = 'Number of unicorns') + 
    theme_grey() + 
    theme(plot.title = element_text(face = "bold")), 
  tooltip = 'text')
```

#### **How long did it take for unicorns to reach their unicorn status from the day they were founded?**

```{r message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=12, fig.height=5, fig.align='center'}
founded_joined <- founded_joined %>% 
  filter(unicorn_time > 0 & unicorn_time < 25)

ggplot(founded_joined, aes(x = unicorn_time, fill = after_stat(count))) + 
  geom_histogram(binwidth = 1, color = 'white') + 
  scale_x_continuous(breaks = seq(0, 40, by = 2)) +
  scale_y_continuous(breaks = seq(0, 200, by = 10)) +
  scale_fill_gradient(low = "cadetblue2", high = "dodgerblue4", name = '') +
  labs(title = 'Time to reach unicorn status',
       x = 'Number of years',
       y = 'Number of unicorns') +
  theme_grey() + 
  theme(plot.title = element_text(face = "bold"),
        legend.position = 'None')
```
